---
title: 'Homework 6: Tessa Anwyll'
output:
  html_document:
    df_print: paged
---


```{r, echo = TRUE}
set.seed(12345)
# Generate random values for y
y <- seq(1, 100, length.out = 1e+08) + rnorm(1e+08)
# Find the mean of y
ybar <- mean(y)
# Initialize variables to use in the for loop 
# to hold sums of squares
sst1 <- 0
sst2 <- 0
# For each element in y, subtract the mean, square 
# it and add it to the current total in the 
# sums of squares vector
system.time({
  for(i in 1:length(y)){
    sst1 <- sst1 + (y[i]-ybar)^2
  }
print(paste("Total Sum of Squares is:", sst1))
})
# Subtract the mean from every element in ybar at once, 
# square each difference then add them up
system.time({
  sst2 <- sum((y-ybar)^2)
  print(paste("Total Sum of Squares is:", sst2))
})
```


Problem 3
```{r, echo = TRUE}
# Given in problem
set.seed(1256)
theta1 <- as.matrix(c(0,0), nrow = 2) #temporarily holds newly created value
theta <- as.matrix(c(1,2), nrow = 2)
X <- cbind(1, rep(1:10, 10))
h <- X %*% theta + rnorm(100, 0, 0.2)
# Initialize variables and parameters
i <- 1 # a counter to have a break point
a <- .0001 # step size
m <- nrow(X) # interval width
tol <- .0000001 # desired difference between successive terms
actualTol <- TRUE
# While you haven't reached tolerance yet, run the loop
while(actualTol == TRUE){
  # Create new values for theta 0 and theta 1
  theta1[1,1] <- theta[1,1] - a*(1/m)*sum(
    theta[2,1]*X[,2]+theta[1,1]-h)
  theta1[2,1] <- theta[2,1] - a*(1/m)*sum((
    theta[2,1]*X[,2]+theta[1,1]-h)*X[,2])
  
  # If you reach desired difference between 
  # successive terms, end the loop
  if((abs(theta1[1,1]-theta[1,1]) > tol) && 
     ((abs(theta1[2,1]-theta[2,1])) > tol)){
    actualTol <- TRUE
  }
  else{
    actualTol <- FALSE
  }
  
  # Store new theta in the location of the old theta
  theta[1,1] <- theta1[1,1]
  theta[2,1] <- theta1[2,1]
  
  # Counter iterates for each run
  i <- i+1
  
  # Stop the loop if you hit 1,000,000 runs
  if(i>1000000){
    actualTol = FALSE
    }
  
}
# Use regression to calculate actual values
act <- lm(h~0+X)

# Create a table with theta estimates, how many loops 
# it took, tolerance, step size and error
outTable <- data.frame(c(theta[1,1], theta[2,1], i, tol, a, 
                         act[[1]][1]-theta[1,1], act[[1]][2]-theta[2,1]))
outTable[,1] <- round(outTable[,1], 5)
rownames(outTable) <- c("Theta 0", "Theta 1", 
                        "Number of Loops", "Tolerance", 
                        "Step Size", "Error theta 0", 
                        "Error theta 1")
colnames(outTable) <- "value"
outTable
```

Problem 4

Rewrite your 
\begin{equation*}
\hat\beta = (X'X)^{-1}X'\underline{y}
\end{equation*} 

as 
\begin{equation*}
(X'X)\hat\beta = X'\underline{y}
\end{equation*} 
and use the solve function to find $\hat{\beta}$
Using solve turns this into a system of equations question which has a faster compute time than an inverted matrix.

Problem 5
```{r data, echo=T, eval=F, include=T}

    set.seed(12456) 
    system.time({
    G <- matrix(sample(c(0,0.5,1),size=16000,replace=T),ncol=10)
    R <- cor(G) # R: 10 * 10 correlation matrix of G
    C <- kronecker(R, diag(1600)) # C is a 16000 * 16000 block diagonal matrix
    id <- sample(1:16000,size=932,replace=F)
    q <- sample(c(0,0.5,1),size=15068,replace=T) # vector of length 15068
    A <- C[id, -id] # matrix of dimension 932 * 15068
    B <- C[-id, -id] # matrix of dimension 15068 * 15068
    p <- runif(932,0,1)
    r <- runif(15068,0,1)
    C<-NULL #save some memory space
   # y <- p - A %*% solve(B)*(q-r)
    })
    print(paste("The size of matrix A is:", object.size(A)))
    print(paste("The size of matrix B is:", object.size(B)))
    
```


*Part A* 

[1] "The size of matrix A is: 112347224 bytes"
[1] "The size of matrix B is: 1816357208 bytes"
Other people ran it and it took about 18 minutes

*Part B*
I just did what I suggested above, I rearranged the equation so there were no more inverses and then used the solve function. I had to multiply the p vector by the I matrix so that it was the same size as B (I'm actually not so sure it's legal though) so that I could right multiply to cancel out the $B^{-1}$. It went slower than the original. 

```{r}
 set.seed(12456) 
system.time({
    G <- matrix(sample(c(0,0.5,1),size=16000,replace=T),ncol=10)
    R <- cor(G) # R: 10 * 10 correlation matrix of G
    C <- kronecker(R, diag(1600)) # C is a 16000 * 16000 block diagonal matrix
    id <- sample(1:16000,size=932,replace=F)
    q <- sample(c(0,0.5,1),size=15068,replace=T) # vector of length 15068
    A <- C[id, -id] # matrix of dimension 932 * 15068
    B <- C[-id, -id] # matrix of dimension 15068 * 15068
    p <- runif(932,0,1)
    r <- runif(15068,0,1)
    C <- NULL #save some memory space
    I <- diag(15068)
    #y <- solve(B, (p*I %*% B)-A*(q-r))
})
```
My output: Timing stopped at: 2547 7.89 2558
 Show Traceback
Error in (p * I %*% B) - A * (q - r) : non-conformable arrays


So obviously my method didn't work very well (at all).