---
title: "Homework 2"
author: "Tessa Anwyll"
date: "August 31, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(tidyr)
```

# Homework 2
*I have neither given nor recieved any unauthorized assistance on this assignment*

## Problem 4
One year in college I had a lengthy, multi-part final project for a class that counted as a large part of my semester grade. I spent a ton of time working on it and had only one part left to do. When I went to finish that one part, the entire rest of the project was gone! I spent an hour digging through my hard drives and trying to recover files to see where it had gone, but I couldn’t find it. I ended up staying up all night until class the next day when it was due re-doing the entire thing. Because it had been on the computer, I had very limited notes to work off of so while I had figured everything out, all the formatting and retyping still took a long time. I wish I known about version control before this, if only for situations like these! Additionally, it seems really easy for collaboration, fixing/finding issues and returning to an old version when your “updates” turn out to be awful and cause everything to break. It seems like it really can provide some peace of mind and increase workflow since you don’t necessarily have to spend hours digging through code to find the one thing that you changed that is causing everything to mess up; you could find it, or just revert to the last version that worked and start back over from there. Plus, if my computer is stolen or destroyed by dropping it, spilling something on it or having some other unfortunate event happen to it, my work will not be lost since it is backed up in many places, not just locally. 


## Problem 5

###A
```{r}
# Read in the data as a list so I can split it up by row
tableA<-(readLines("https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"))

# Get information about the data
tableA<-strsplit(tableA, split = " ")

matA<-tableA[3:length(tableA)]
dfnames<-vector()
for(i in 1:length(matA)){
  
  if(length(matA[[i]])==6){
   dfnames<-c(dfnames, paste("Item ",matA[[i]][1]))
  matA[[i]]<-c(matA[[i]][2:6])

  }
} 


aDf<- matrix(unlist(matA), nrow = 10, byrow = TRUE, ncol=5)

rownames(aDf)<-dfnames
colnames(aDf)<-c("Operator 1", "Operator 2", "Operator 3", "Operator 4", "Operator 5")
aDf<-data.frame(aDf)
#aDf<-gather(aDf, key = Operator, value = c(1, 2, 3, 4, 5), -Operator.1, -Operator.2, -Operator.3, -Operator.4, -Operator.5)
aDf
```

###B
```{r}
# Read in the data
tableB<-read.csv("http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat")

# Get information about the data
str(tableB)

# Look at the data
tableB
```
```{r, echo = TRUE}
## Separate the data into columns
tableB1<-separate(tableB, Year.Long.Jump.Year.Long.Jump.Year.Long.Jump.Year.Long.Jump, into = c("Y1", "LJ1", "Y2", "LJ2", "Y3", "LJ3", "Y4", "LJ4"), sep = " ")

## Consolidate all year and long jump values into two variables and filter out NA 
year<-as.vector(as.numeric(c(tableB1[,1], tableB1[,3], tableB1[,5], tableB1[,7])))
LongJump<-as.vector(as.numeric(c(tableB1[,2], tableB1[,4], tableB1[,6], tableB1[,8])))
matB1<-cbind(tableB1, year, LongJump)
tableB1<-select(matB1, year, LongJump)
tableB1<-filter(tableB1, !is.na(tableB1$year)&!is.na(tableB1$LongJump))

## Convert years since 1900 into years 
tableB1<-mutate(tableB1, Year = year+1900)
tableB1<-select(tableB1, Year, LongJump)
colnames(tableB1)<- c("Years (years)", "Long Jump (in)")
tableB1

```


###C
```{r}
dataC<-fread("https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat")

colnames(dataC)<-c("BodyWeight1", "BrainWeight1", "BodyWeight2", "BrainWeight2", "BodyWeight3", "BrainWeight3", "BodyWeight4", "BrainWeight4", "BodyWeight5", "BrainWeight5", "BodyWeight6", "BrainWeight6")

BoWe<-vector()
BrWe<-vector()
i<-1
for(i in 1:3){
 BoWe<-c(BoWe, dataC[,(paste("BodyWeight",i))])
 BrWe<-c(BrWe, dataC[,string(paste("BrainWeight",i))])
}
BoWe
BrWe
#for(i in 1:12){
#BoWe <- c(BoWe, dataC[ ,!i%%2])
#BrWe<- c(BrWe, dataC[ ,i%%2])
#}
#dataC<-cbind(dataC, BoWe, BrWe)
#for(j in 1:12){
 # dataC[,j]
#}

#dataC<- mutate(dataC, BodyWeight = BoWe, BrainWeight = BrWe)
#dataC<-select(dataC, BodyWeight, BrainWeight)
#dataC
```


## Problem 6

Import the data and take a look at what we are working with:
```{r include=FALSE}
# Path to data
library(swirl)
.datapath <- file.path(path.package('swirl'), 'Courses',
                      'R_Programming_E', 'Looking_at_Data',
                      'plant-data.txt')
# Read in data
plants <- read.csv(.datapath, strip.white=TRUE, na.strings="")

# Remove annoying columns
.cols2rm <- c('Accepted.Symbol', 'Synonym.Symbol')
plants <- plants[, !(names(plants) %in% .cols2rm)]

# Make names pretty
names(plants) <- c('Scientific_Name', 'Duration', 'Active_Growth_Period',
                   'Foliage_Color', 'pH_Min', 'pH_Max',
                   'Precip_Min', 'Precip_Max',
                   'Shade_Tolerance', 'Temp_Min_F')

```
```{r echo=TRUE}
# Look at the data
plants
```

The biggest issue is all the "NA"s. My next move is to remove each row that does not have NAs in any of my variables I will be testing (so no NAs for Foliage Color or either of the pH measures)


```{r}
library(dplyr)
# Store new data frame to use in my test with only rows with no "NA"s for my variables of interest
newplants<-filter(plants,!is.na(plants$pH_Min)&!is.na(plants$pH_Max)&!is.na(plants$Foliage_Color))

# Create a new variable that takes the average of the min and max pH reported in the data set
newplants<-mutate(newplants, pHAvg = (pH_Min+pH_Max)/2)

# Group by foliage color, my predictor
newplants<-group_by(newplants, Foliage_Color)

# Take a look at the data set now
newplants

# Make a table just to display that summarizes the means of each average ph and counts for each foliage color
dispnewplants <- summarize(newplants, n(), mean(pHAvg))
dispnewplants

```
Now that the data is ready to go, I will run my ANOVA test and interpret my results at a .05 significance level
```{r, echo = TRUE}

# Store regression model in a new variable
ANOVAtestdata<-lm(pHAvg~Foliage_Color, newplants)

# Run ANOVA Test
ANOVAResults<-anova(ANOVAtestdata)

# Format output as a nicer table
library(magrittr)
library(kableExtra)
ANOVAResults %>%
  kable() %>%
  kable_styling(full_width = TRUE, bootstrap_options = c("striped", "condensed"))

```

With a p value (0.0015) so much lower than my significance level (.05), we can concluded that there is a statistically significant relationshipe between foliage color and pH.
**coefficients?**


